---
title: "Google Civic vs. WI Election Commission Polling Places"
subtitle: 
author:
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)

library(httr)
library(jsonlite)
library(tidyverse)
library(magrittr)
library(here)
library(knitr)
library(kableExtra)
library(rvest)

kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```


# Polling locations via Google Civic's API

(Actually, it goes State (elections.wi.gov) --> Voter Information Project --> Google Civic's API)

To check the state's data's accuracy, we look up polling locations for addresses in every voting precinct using the Google Civic API.

```{r}
#########################################################
# one address per precinct
addresses <- read_csv(here::here("data", "wisconsin_example_addresses.csv"))

addresses %<>% mutate_if(is.character, str_to_upper)

addresses %<>% arrange(county_name)
```



Polling locations are sometimes (in the days or weeks before an election) available via the Google Civic API, which uses data collected from states by the Voter Information Project, which notifies users when it posts data here: https://groups.google.com/g/vip-community

Notes about the Google Civic API (see the relevant documentation [here](https://developers.google.com/civic-information/docs/v2/elections/voterInfoQuery))  
- create an API key here: https://console.developers.google.com/apis/credentials  
- remember to authorize you API key  
- API calls should not need `electionId`; docs say it is optional  
- `pollingLocations` will not always be returned  
- may return a great deal more than polling locations (see Additional Google Civic API Contents below)
- may return more than one polling place per input address as a list 



```{r}
# make full street address
addresses %<>% 
  mutate(voter_address = paste(voting_street_address,
                               voting_city, 
                               state_code))

addresses$voter_address[2]
```

Here is the polling place from the Google Civic API for the above address:
```{r}

# need to get AND authorize an api key
source(here::here("api_key.R"))

# A function to get poll addresses
get_poll <- function(voter_address){
  
  voter_address %<>% 
    str_replace_all(" ", "%20") %>%
    str_remove_all(",")
  
  url <- str_c("https://civicinfo.googleapis.com/civicinfo/v2/voterinfo?address=", 
               voter_address, 
               "&returnAllAvailableData=true&key=", 
               api_key)
  
  
  poll <- GET(url) %>% 
    .$content %>% 
    rawToChar() %>% 
    fromJSON() %>%
    .$pollingLocations %>% # note: there is a lot more than this
    #.$address %>% 
    {ifelse(is.null(.), "NULL", .)} %>% 
    unlist() %>% 
    str_c(collapse = ", ")
  
  return(poll)
}

get_poll(addresses$voter_address[2])
```

```{r, eval = FALSE}
# apply function to addresses
api_locations <- addresses %>% 
  # "NULL" if NULL, "error" if any error
  mutate(GC_Polling_Place = voter_address %>% map_chr(possibly(get_poll, "error"))) 

## For some reason GC returnds 400 or 500 observations and a lot of nulls each pass--does not appear to be a rate limit thing, so we just do pass after pass until we are not getting any more. 

# initialize found 
api_found <- api_locations %>% filter(!GC_Polling_Place %in% c("error", "NULL"))

# initialize missing
api_missing <- api_locations %>% filter(GC_Polling_Place %in% c("error", "NULL"))

##############################################
# loop until api_found approaches api_locations
n = 0
while(n < 5 | nrow(api_missing) < nrow(api_locations_temp) ){  
  
api_locations_temp <- api_missing %>% 
  # "NULL" if NULL, "error" if any error
  mutate(GC_Polling_Place = voter_address %>% map_chr(possibly(get_poll, "error"))) 

api_missing <- api_locations_temp %>% filter(GC_Polling_Place %in% c("error", "NULL"))

api_found %<>% full_join(api_locations_temp %>% 
                           filter(!GC_Polling_Place %in% c("error", "NULL")))
message(paste( 
  nrow(api_locations_temp) - nrow(api_missing), "new,",
  nrow(api_found), "found in all,",
  nrow(api_missing), "left."))

# run 5 times if 0 new
if(nrow(api_missing) == nrow(api_locations_temp)){n = n + 1}
   
}
# END LOOP 
#########################################

GC_api_locations <- addresses %>% left_join(api_found) %>% select(-starts_with("vot"))

# csv for github
write_csv(GC_api_locations, here("data", "GC_api_locations.csv"))

# google sheet 
library(googlesheets4)

ss <- googledrive::drive_get("WI polling locations") %>%
  gs4_get() 

GC_api_locations %>%
  sheet_write(ss = ss,
              sheet = paste("GC Polling Locations", Sys.Date() ))
```

### Sample API results 

```{r}
api_locations <- read_csv(here("data", "GC_api_locations.csv"))

# summary table
api_locations %>% 
  head() %>% 
  select(precinct_name, precinct_id, GC_Polling_Place) %>%  
  kablebox()
```

# Google Civic API results compared to the WEC file from 10-27

```{r}
# from polling wards rmd
upload <- read_csv(here::here("data", "VAN Polling Locations 2020-10-21.xlsx"))

names(upload) %<>% str_to_lower() %>% str_replace(" ", "_")

upload %<>%
  mutate(address = address %>% 
           str_replace(" RD", " ROAD") %>%
           str_replace("MT\\.", "MOUNT") %>% 
           str_squish(),
         polling_location = polling_location %>% 
           str_replace(" RD", " ROAD") %>%
           str_replace("MT\\.", "MOUNT") %>% 
           str_squish())

api_locations <- read_csv(here("data", "GC_api_locations.csv"))

names(api_locations) %<>% str_to_lower() %>% str_replace(" ", "_")

api_locations %<>%
  mutate_if(is.character, str_to_upper) %>% 
  mutate(gc_polling_location = gc_polling_place %>% 
           str_replace(" RD", " ROAD") %>%
           str_replace("MT\\.", "MOUNT") %>% 
           str_squish() )

# join on precinct id and name 
d <- full_join(upload, api_locations)

d %<>% select(precinct_id, 
              precinct_name,
              polling_location,
              address,
              zip,
              gc_polling_location)
```

Potentially different zip
```{r}
zips <- d %>% filter(!str_detect(gc_polling_location, as.character(zip)))

zips %>% kablebox()
```

Potentially different address
```{r}
adds <- d %>% filter(!str_detect(gc_polling_location, address))

adds %>% kablebox()
```

Potentially different polling location names
```{r}
locs <- d %>% filter(!str_detect(gc_polling_location, polling_location)) # %>% select(polling_location, gc_polling_location) %>% distinct()

locs %>% kablebox()
```

## Diff with WEC [here](https://docs.google.com/spreadsheets/d/1cFJX0yGY91bzDpaR9ODwCeSCbzUrkWpRb1m7HwMKOA4/edit#gid=240141679)

```{r, eval = FALSE}
gc_diff <- full_join(zips, adds) %>% full_join(locs) %>% 
  # drop missing
  filter(!is.na(zip))

# csv for github
write_csv(gc_diff, here("data", "GC_api_WEC_diff.csv"))

# google sheet 
library(googlesheets4)

ss <- googledrive::drive_get("WI polling locations") %>%
  gs4_get() 

gc_diff %>%
  sheet_write(ss = ss,
              sheet = paste("GC issues", Sys.Date() ))
```

# Additional Google Civic API Contents

For reference, these values may or may not be returned: 
```
"pollingLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "earlyVoteSites": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "dropOffLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
```