---
title: "Changes in Polling Places"
subtitle: 
author:
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)

library(httr)
library(jsonlite)
library(tidyverse)
library(magrittr)
library(here)
library(knitr)
library(kableExtra)
library(rvest)

kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```


## TL;DR: Download VAN-formatted elections.wi.gov + Google Civic polling locations [here](https://docs.google.com/spreadsheets/d/1cFJX0yGY91bzDpaR9ODwCeSCbzUrkWpRb1m7HwMKOA4/edit?usp=sharing)


This link will update when the WEC file and date in the code chunk below are updated.

# Polling Locations from the state 

We are currently using this file:

```{r}
# UPDATE THIS IF A NEW FILE IS POSTED! 
file <- "Polling Place Locations - 2020 General Election Updated 10-21-2020.xlsx"

`Date Authenticated*` <- "2020-10-21"

upload_file <- paste0("VAN Polling Locations ", `Date Authenticated*`, ".xlsx")

tibble(file,
       `Date Authenticated*`) %>% 
  kablebox()

official_locations <- readxl::read_xlsx(here::here("data", file))

official_locations %<>% mutate_if(is.character, str_to_upper) %>%
  mutate(PollingPlaceAddress = PollingPlaceAddress %>% 
           str_replace(" RD", " ROAD"),
         County = str_remove(County, "COUNTY"))

# some locations have more than one room and some munis have duplicates
official_locations %<>% select(-VotingRoomArea)

# combine polling places at the same address
official_locations %<>%
  group_by(County, PollingPlaceAddress, PollingPlaceName, Muni) %>%
  mutate(ReportingUnit = paste(unique(County), 
                               unique(Muni), 
                               unique(ReportingUnit), 
                               sep = " - ",
                               collapse = " and ")) %>%
  ungroup() %>%
  distinct()
```

The most recent xlsx file of polling locations that is posted: https://elections.wi.gov/node/6527

```{r}
# polling locations posted on website
html <- read_html("https://elections.wi.gov/node/6527") %>%
  html_nodes("a") 

files <- tibble(file = html_text(html),
                link = html_attr(html, "href")) %>% 
  filter(str_detect(file, "xls|csv")) 
                
kablebox(files)
```

---

# Polling locations via Google Civic's API

(Actually, it goes State (elections.wi.gov) --> Voter Information Project --> Google Civic's API)

To check the state's data's accuracy, we look up polling locations for addresses in every voting precinct using the Google Civic API.

```{r}
#########################################################
# one address per precinct
addresses <- read_csv(here::here("data", "wisconsin_example_addresses.csv"))

addresses %<>% mutate_if(is.character, str_to_upper)

addresses %<>% arrange(county_name)

# FIXME DELETE THIS WHEN API IS WORKING
addresses %<>% head(2)
# /FIXME
```



Polling locations are sometimes (in the days or weeks before an election) available via the Google Civic API, which uses data collected from states by the Voter Information Project, which notifies users when it posts data here: https://groups.google.com/g/vip-community

Notes about the Google Civic API (see the relevant documentation [here](https://developers.google.com/civic-information/docs/v2/elections/voterInfoQuery))  
- create an API key here: https://console.developers.google.com/apis/credentials  
- remember to authorize you API key  
- API calls should not need `electionId`; docs say it is optional  
- `pollingLocations` will not always be returned  
- may return a great deal more than polling locations (see Additional Google Civic API Contents below)
- may return more than one polling place per input address as a list 



```{r}
# make full street address
addresses %<>% 
  mutate(voter_address = paste(voting_street_address,
                               voting_city, 
                               state_code))

addresses$voter_address[1]
```

Here is the polling place from the Google Civic API for the above address:
```{r}

# need to get AND authorize an api key
source(here::here("api_key.R"))

# A function to get poll addresses
get_poll <- function(voter_address){
  
  voter_address %<>% 
    str_replace_all(" ", "%20") %>%
    str_remove_all(",")
  
  url <- str_c("https://civicinfo.googleapis.com/civicinfo/v2/voterinfo?address=", 
               voter_address, 
               "&returnAllAvailableData=true&key=", 
               api_key)
  
  
  poll <- GET(url) %>% 
    .$content %>% 
    rawToChar() %>% 
    fromJSON() %>%
    .$pollingLocations %>% # note: there is a lot more than this
    .$address %>% 
    {ifelse(is.null(.), "NULL", .)}
  
  return(poll)
}


# test 
get_poll(addresses$voter_address[1])

# apply function to addresses
api_locations <- addresses %>% 
  # "NULL" if NULL, "error" if any error
  mutate(PollingPlaceAddress = voter_address %>% map(possibly(get_poll, "error"))) 

# expand if there is more then one polling place per address
api_locations %<>% unnest(PollingPlaceAddress)

api_locations %>% select(voting_street_address, precinct_name, precinct_id, PollingPlaceAddress) %>%  kablebox()
```

### API results not in the state's xlsx file

```{r}
api_locations %>% 
  filter(PollingPlaceAddress != "NULL") %>% # filter out API errors
  anti_join(official_locations) %>% # filter out locations already posted
  select(precinct_name, precinct_id, PollingPlaceAddress) %>%
  kablebox()
```

### Cases where Google Civic's API did not return the same address when given polling locations

> Until the API is working, it is all of them--just showing a few for now.

```{r}
official_locations %>% 
  head() %>% #FIXME WHEN API IS WORKING
  mutate(PollingPlaceAddress_API = map(PollingPlaceAddress, get_poll)) %>% # get polling places for polling places' addresses
  filter(PollingPlaceAddress_API != PollingPlaceAddress) %>% # mismatches
  select(Muni, ReportingUnit, PollingPlaceName, PollingPlaceAddress, PollingPlaceAddress_API) %>%
  kablebox()
```

### Polling place addresses not returned by Google Civic's API

> Until the API is working, it is all of them--just showing a few for now.

```{r}
official_locations %>% 
  head() %>% #FIXME WHEN API IS WORKING
  anti_join(api_locations) %>% # remove ones matched with API
  select(Muni, ReportingUnit, PollingPlaceName, PollingPlaceAddress) %>% 
  kablebox()
```

---

## Combine results

```{r}
# name both official and API results to fit VAN
api_locations %<>% rename(County = county_name)

# reformat county
official_locations %<>% 
  mutate_if(is.character, str_to_upper) %>% 
  mutate(County = str_remove(County, " COUNTY") )


locations <- official_locations

locations %<>% full_join(api_locations)
```

## Format for VAN upload 

Notes: 

- Several VAN variable names have spaces in them 
- VAN has one row per ward, but the WEC file may say Wards 1-5
- VAN addresses are sometimes title case and sometimes all caps. I am making all addresses all caps for the best matching prospects. (WEC polling locations are in all caps.)



Here are the top few rows of the VAN file from 8-11-2020, (the August 2020 Primary):

```{r}
# old VAN data
van <- readxl::read_xlsx(here("data", "VAN Polling Locations.xlsx"))

# format strings consistantly as upper case
van %<>% 
  mutate_if(is.character, str_to_upper) %>%
  mutate(Address = Address %>% 
           str_replace(" RD", " ROAD"))

van %>% head() %>% kablebox()

# name combined official and API results to fit VAN
van_new <- locations %>% 
  mutate(Address = str_remove(PollingPlaceAddress, ", .*") %>% 
           str_to_upper(),
         County = str_to_upper(County) %>% 
           str_remove(" COUNTY"),
         `Zip` = str_extract(PollingPlaceAddress,"WI [0-9]*") %>% 
           str_remove("WI "),
         # Muni, Phone, and Location Name is from the state
         `Polling Location` = str_to_upper(PollingPlaceName),# %>% str_remove("^School "), #FIXME if we are altering names in the state data to match the VAN data 
         `Polling Location Contact Info` = Phone,
         # Precinct ID  and name is from API
         `Precinct ID` = precinct_id,
         `Precinct Name` = precinct_name,
         `Date Authenticated*` = `Date Authenticated*`,
         `Authenticating Authority` = ifelse(is.na(ReportingUnit), "Google Civic", "State Election Official") )

# format as compact strings 
van_new %<>% 
  mutate_all(as.character) %>%
  mutate_all(str_squish)

van_new$`Precinct ID` %<>% as.numeric()

# a function to split out wards
split <- function(ward){
  
  min = str_extract(ward, "[0-9]*") %>% 
           as.numeric()
  
  max = str_extract(ward, "-[0-9]*") %>% 
           as.numeric() %>% abs()
  
  full_seq(c(min,
             max),
           1) %>%
    str_c(collapse = ",") 
}

# split out wards
van_new %<>% #select(ReportingUnit) %>% 
  # undo earlier merging
  group_by(PollingPlaceAddress, PollingPlaceName) %>% 
  mutate(ReportingUnit = str_split(ReportingUnit, " AND ") ) %>% 
  unnest(ReportingUnit) %>% 
  mutate(ward = str_extract(ReportingUnit, "[1-9].*") %>% 
           # split on commas in wards
           str_split(",")) %>% 
  unnest(ward) %>% 
  mutate(ward_seq = map_chr(ward, possibly(split, otherwise = NA)),
         ward = coalesce(ward_seq, ward) %>% 
           str_split(",")) %>% 
  unnest(ward) %>% 
  mutate(ward = ward %>% str_pad(3, side = "left", pad = "0")) %>% 
  mutate(ReportingUnit = str_replace(ReportingUnit, " WARD(|S) .*", paste(" WARD", ward))) %>% 
  ungroup() %>% 
  select(-ward, -ward_seq) %>%
  select(ReportingUnit, everything()) %>% 
  distinct() %>% 
  add_count(ReportingUnit) %>% 
  arrange(n, ReportingUnit) # %>% filter(n >1)



# make sure all required columns not coming from VAN are there
tibble(VAN_column = names(van),
       exists_in_new_data = names(van) %in% names(van_new)) %>% 
  kablebox()
```

# Discrepancies between WEC and VAN data

## Join with old VAN upload data

```{r}
upload <- van_new %>%
  filter(Address != "NULL") %>% 
  select(PollingPlaceAddress,
         #Muni, 
         ReportingUnit, 
         any_of(names(van)) ) %>% 
  left_join(van %>% 
              # drop info that we want to replace, never retain 
              select(-`Polling Location Contact Info`, 
                     -`Authority Contact Info`,
                     -`Date Authenticated*`,
                     -`Authenticating Authority`),
            # NOTE: requiring exact match on polling location name, which can cause mismatches
            by= c("County", 
                  "Address", 
                  #"City", 
                  "Zip")) %>% 
  distinct()
```

### Discrepancies in Polling Location Names

> NOTE: We are not matching on the polling location name, so this is just a check for false positives or name changes that may have occurred without an address change.

#### Cases where the state's polling location name (x) is not exactly the same as the last VAN upload's polling location name (y):

```{r}
# where polling location strings are not the same
upload %>% 
  filter(`Polling Location.x` != `Polling Location.y` ) %>% 
  select(`Polling Location.x`,  
         `Polling Location.y`, 
         Address, 
         City, 
         Zip, County) %>% 
  distinct() %>% 
  kablebox()
```


> Precinct Names and IDs should match, but they are not from the State's file, so if google civic fails to return a matching polling place address for the example address in each precinct, Precinct Names and IDs will be `NA`.

> We are not matching on Precinct Name or ID. Where Precinct Names differ between the example addresses and VAN, we use the Precinct Name from the example addresses (but we could do the reverse).


```{r}
upload %<>% 
  # take the State's name first 
  mutate(`Precinct ID` = coalesce(`Precinct ID.x`,
                                  `Precinct ID.y`,),
         `Precinct Name` = coalesce(`Precinct Name.x`,
                                    `Precinct Name.y`),
         `Polling Location` = coalesce(`Polling Location.x` , 
                                       `Polling Location.y`)
         ) %>% 
  # select variable names from van
  select(any_of(names(van)), ReportingUnit,  PollingPlaceAddress) %>% 
  distinct()
```

### VAN polling place addresses not in WEC's file

#### Cases where the polling location address that is in VAN is not matching in WEC file

These polling places may have been deleted
```{r}
# polling addresses not in VAN? 
deleted <- van %>%  filter(!Address %in% upload$Address) %>% 
  select(County,  Address, `Polling Location`, `Precinct Name`, `Date Authenticated*`) %>% 
  distinct() 

deleted %>% kablebox()
```

### WEC polling place addresses not in VAN

#### Cases where the polling location address is not matching in VAN 

These may be new polling places
```{r}
# polling addresses not in VAN? 
added <- upload %>% 
  filter(is.na(`Precinct Code`),
         !Address %in% van$Address) %>% 
  select(County, ReportingUnit,  `Precinct Name`,Address, `Polling Location`, `Date Authenticated*`) %>% 
  distinct() 

added %>% kablebox()

deleted %<>% mutate(Issue = "Deleted address")
added %<>% mutate(Issue ="New address")

poll_address_changes <- full_join(added, deleted)  %>% select(Issue, everything())

poll_address_changes
```

Potentially added or deleted polling locations are automatically created [here](https://docs.google.com/spreadsheets/d/1cFJX0yGY91bzDpaR9ODwCeSCbzUrkWpRb1m7HwMKOA4/edit#gid=740092819). When we confirm that these are actual changes and not erroneous mismatches, they should be added to the [Polling Location Issue Tracker](https://docs.google.com/spreadsheets/d/1GYjifHWJ-N_Z3YEHMVWp58lA5d_emrzvgd4O951UTdU/edit#gid=120487606)

```{r, eval=FALSE}
library(googlesheets4)
#FIXME pull previously flagged issues and put new ones at the top

ss <- googledrive::drive_get("polls") %>%
  gs4_get() 

poll_address_changes %>%
  sheet_write(ss = ss,
              sheet = str_c(Sys.Date() %>% as.character(), 
                            " Issues"))
```

#### Cases where the county + address is not matching in VAN 

These may be "ghost wards" but still may be worth checking.

```{r}
# polling addresses + counties not in VAN? 
upload %>% 
  filter(is.na(`Precinct Code`),
         !Address %in% poll_address_changes$Address) %>% 
  select(County,  PollingPlaceAddress, `Polling Location`) %>% 
  distinct() %>% 
  kablebox()
```

---

These could be new locations or something else, like how the State says that people in Bayfield Ward 12 are now going into Ashland to vote???

```{r}
# last VAN upload 
van %>% 
  filter(str_detect(Address, "BEASER")) %>% 
  select(County, `Precinct Name`, Address) %>% distinct() %>% kablebox()

# Current state data
official_locations %>%
  filter(str_detect(PollingPlaceAddress, "BEASER")) %>% 
  select(County,  Muni, ReportingUnit, PollingPlaceAddress) %>% kablebox()

```

### All locations being updated in any way

(including location name changes and new addresses)

In old VAN, not in upload:
```{r}
anti_join(van, upload, by = c("County", "Address", "Polling Location")) %>% kablebox()
```

In new upload, not in old VAN:
```{r}
anti_join(upload, van, by = c("County", "Address", "Polling Location")) %>% kablebox()
```

# Save new polling location data 

```{r, eval=FALSE}
write_csv(upload, path = here("data", upload_file))

library(googlesheets4)

ss <- googledrive::drive_get("WI polling locations") %>%
  gs4_get() 

upload %>%
  sheet_write(ss = ss,
              sheet = paste("VAN Polling Locations", `Date Authenticated*`))
```

Here are the first few rows of the upload:
```{r}
upload %>% head() %>% kablebox()
```


## VAN-formatted combined elections.wi.gov + Google Civic polling locations:

CSV file: `r paste0("https://raw.githubusercontent.com/judgelord/vote/master/data/", upload_file) %>% str_replace_all(" ", "%20")`

Google Sheet: https://docs.google.com/spreadsheets/d/1cFJX0yGY91bzDpaR9ODwCeSCbzUrkWpRb1m7HwMKOA4

---

# Q & A
- Which fields from the VAN must be replaced exactly? If the state changes how `Polling Location` is described, do we want to replace it or have it fail to match with the `Precinct ID` from the VAN template? There are over 200 polling locations with slightly different names in the file on the state's website and the old VAN data.

>Precinct ID, Precinct Code, and Precinct Name are ones that should always be the same from VAN. If the state's name is slightly different for a polling location, if it's easier to just change it to that, that is fine. If the polling location is described differently, just replace it with what the state has. I think that answers your question, but let me know if it doesn't.

- There are over 200 polling location addresses in the State's 8-11-2020 xlsx file that I do not see in the VAN list. I only investigated the first one, 2100 Beaser Ave, Ashland, WI, and I am not clear if Bayfield Ward 12 is now voting there or not. Is there a simple explanation? If they need to be investigated, Can I anything with the data to make these cases easier to investigate? 

>These are most likely what we call ghost wards, where no one lives in. So we just exclude them from uploading to VAN since they don't have a VAN Precinct ID.

- The state's `Muni` is not the same as `City` in the VAN data for almost 2000 polling place addresses. For now, `Muni` is not being used in the merge. 
- If we are trying to align `City` and `Muni`, `City` in the VAN data does not contain "Town of"/"City of"/"Village of" but both the state's poll locations `Muni` xlsx file and the addresses file do. We are dropping the epithet, correct?

>This is probably because, in smaller munis, the actual address for a place has another muni in it. An example is the Town of Adams in Adams County. The actual address for the Town of Adams Town Hall has Friendship as the muni. So I would continue not using muni to match. We don't need to align the actual municipality name with the address municipality name. 

- If Google Civic has polling locations that the file on the state website does not, what should `Authenticating Authority` be? It is "Google Civic API" for now. 
- `Authority Contact Info` is "DPW- Fall Partisan Primary" in the VAN file. Do we want this to take the new file name or WEC's phone number?

>This is something that gets applied in bulk after I upload polling locations. So if it's easier for you to just not deal with those, that's fine. If it isn't, then yeah, Google Civic can be the authenticating authority and DPW-November Election as contact info

- `Location Description` in the VAN file contains poll hours. Do we want to merge this into the new upload or try to pull it in from Google Civic? Or both? Or neither?

> Keep location description as is with the polling hours. 

- I assume we are merging in `Precinct Code` from the VAN file on `Precinct ID`, correct? 

>Precinct ID is the primary key for uploading into VAN


- If a polling location is removed (i.e., it was in the previous VAN upload), but it is not in the new data, do we keep it in the upload sheet with some kind of a note, or do we just note on this data merge page that locations were removed?

> Don't need it noted in the file to upload but noted in the data merge page works. 

# TODO 

- split out reporting units from the state to see if wards move addresses (not just if there are new or deleted addresses)

# Notes about data for VAN

1. The Addresses in the VAN data capitalize oddly--"1St" "2Nd" "3Rd" etc.--I made all strings all caps.

# Additional Google Civic API Contents

For reference, these values may or may not be returned: 
```
"pollingLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "earlyVoteSites": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "dropOffLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
```