---
title: "Changes in Polling Places"
subtitle: 
author:
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)

library(httr)
library(jsonlite)
library(tidyverse)
library(magrittr)
library(here)
library(knitr)
library(kableExtra)
library(rvest)

kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```


## TL;DR: Download VAN-formatted elections.wi.gov + Google Civic polling locations [here](https://docs.google.com/spreadsheets/d/1cFJX0yGY91bzDpaR9ODwCeSCbzUrkWpRb1m7HwMKOA4/edit?usp=sharing)


This link will update when the WEC file and date in the code chunk below are updated.

# Polling Locations from the state 

We are currently using this file:

```{r}
# UPDATE THIS IF A NEW FILE IS POSTED! 
file <- "Polling Place Locations - 2020 General Election Updated 10-21-2020.xlsx"

`Date Authenticated*` <- "2020-10-21"

upload_file <- paste0("VAN Polling Locations ", `Date Authenticated*`, ".xlsx")

tibble(file,
       `Date Authenticated*`) %>% 
  kablebox()

official_locations <- readxl::read_xlsx(here::here("data", file))

official_locations %<>% mutate_if(is.character, str_to_upper) %>%
  mutate(PollingPlaceAddress = PollingPlaceAddress %>% 
           str_replace(" RD", " ROAD") %>%
           str_replace("MT\\.", "MOUNT"),
         County = str_remove(County, "COUNTY"))

# some locations have more than one room and some munis have duplicates
official_locations %<>% select(-VotingRoomArea)

# combine polling places at the same address
official_locations %<>%
  group_by(County, PollingPlaceAddress, PollingPlaceName, Muni) %>%
  mutate(ReportingUnit = paste(unique(County), 
                               unique(Muni), 
                               unique(ReportingUnit), 
                               sep = " - ",
                               collapse = " and ") %>%
           str_replace("MT\\.", "MOUNT")) %>%
  ungroup() %>%
  distinct()
```

The most recent xlsx file of polling locations that is posted at https://elections.wi.gov/node/6527

```{r}
# polling locations posted on website
html <- read_html("https://elections.wi.gov/node/6527") %>%
  html_nodes("a") 

files <- tibble(file = html_text(html),
                link = html_attr(html, "href")) %>% 
  filter(str_detect(file, "xls|csv")) 
                
kablebox(files)
```

---

## Combine results

```{r}

# reformat county
official_locations %<>% 
  mutate_if(is.character, str_to_upper) %>% 
  mutate(County = str_remove(County, " COUNTY") )


locations <- official_locations
```

## Format for VAN upload 

Notes: 

- Several VAN variable names have spaces in them 
- VAN has one row per ward, but the WEC file may say Wards 1-5
- VAN addresses are sometimes title case and sometimes all caps. I am making all addresses all caps for the best matching prospects. (WEC polling locations are in all caps.)



Here are the top few rows of the VAN file from 8-11-2020, (the August 2020 Primary):

```{r}
# old VAN data
van <- readxl::read_xlsx(here("data", "VAN Polling Locations.xlsx"))

# format strings consistantly as upper case
van %<>% 
  mutate_if(is.character, str_to_upper) %>%
  mutate(Address = Address %>% 
           str_replace(" RD", " ROAD"))

van %>% head() %>% kablebox()

# name combined official and API results to fit VAN
van_new <- locations %>% 
  mutate(Address = str_remove(PollingPlaceAddress, ", .*") %>% 
           str_to_upper(),
         County = str_to_upper(County) %>% 
           str_remove(" COUNTY"),
         `Zip` = str_extract(PollingPlaceAddress,"WI [0-9]*") %>% 
           str_remove("WI "),
         # Muni, Phone, and Location Name is from the state
         `Polling Location` = str_to_upper(PollingPlaceName),# %>% str_remove("^School "), #FIXME if we are altering names in the state data to match the VAN data 
         `Polling Location Contact Info` = Phone,
         `Date Authenticated*` = `Date Authenticated*`,
         `Authenticating Authority` = ifelse(is.na(ReportingUnit), "Google Civic", "State Election Official") )

# format as compact strings 
van_new %<>% 
  mutate_all(as.character) %>%
  mutate_all(str_squish)


# a function to split out wards
split <- function(ward){
  
  min = str_extract(ward, "[0-9]*") %>% 
           as.numeric()
  
  max = str_extract(ward, "-[0-9]*") %>% 
           as.numeric() %>% abs()
  
  full_seq(c(min,
             max),
           1) %>%
    str_c(collapse = ",") 
}

# split out wards
van_new %<>% #select(ReportingUnit) %>% 
  # undo earlier merging
  group_by(PollingPlaceAddress, PollingPlaceName) %>% 
  mutate(ReportingUnit = str_split(ReportingUnit, " AND ") ) %>% 
  unnest(ReportingUnit) %>% 
  mutate(ward = str_extract(ReportingUnit, "[1-9].*") %>% 
           # split on commas in wards
           str_split(",")) %>% 
  unnest(ward) %>% 
  mutate(ward_seq = map_chr(ward, possibly(split, otherwise = NA)),
         ward = coalesce(ward_seq, ward) %>% 
           str_split(",")) %>% 
  unnest(ward) %>% 
  mutate(ward = ward %>% 
           str_remove("[A-z]") %>% 
           str_pad(3, side = "left", pad = "0")) %>% 
  mutate(ReportingUnit = str_replace(ReportingUnit, " WARD(|S) .*", paste(" WARD", ward))) %>% 
  ungroup() %>% 
  select(-ward_seq) %>%
  select(ReportingUnit, everything()) %>% 
  distinct() %>% 
  add_count(ReportingUnit) %>% 
  arrange(n, ReportingUnit) # %>% filter(n >1)



# make sure all required columns not coming from VAN are there
tibble(VAN_column = names(van),
       exists_in_new_data = names(van) %in% names(van_new)) %>% 
  kablebox()
```

# Discrepancies between WEC and VAN data

## Join with old VAN upload data

```{r}
# add in precinct IDs for new addressses
van_new %<>% 
  rename(`Precinct Name` = ReportingUnit) %>%
  left_join(van %>% select(`Precinct ID`, 
                           `Precinct Code`, 
                           `Precinct Name`,
                           City)) 

#FIXME inspect NAs
van_new %<>% 
  select(`Precinct Name`, `Precinct ID`, everything()) %>%
  distinct()

upload <- van_new %>%
  filter(Address != "NULL") %>% 
  select(PollingPlaceAddress,
         any_of(names(van)) ) %>% 
  left_join(van %>% 
              # drop info that we want to replace, never retain 
              select(-`Polling Location Contact Info`, 
                     -`Authority Contact Info`,
                     -`Date Authenticated*`,
                     -`Authenticating Authority`,),
            # NOTE: requiring exact match on polling location name can cause mismatches
            by= c("Precinct Name", "Precinct Code", "Precinct ID", "County", "City", "Zip"),
            suffix = c("_WEC", "_VAN")) %>% 
  distinct()


```

### Discrepancies in Polling Location Names

> NOTE: We are not matching on the polling location name, so this is just a check for false positives or name changes that may have occurred without an address change.

#### Cases where the addresses match, but the state's polling location name is not exactly the same as the last VAN upload's polling location name:

```{r}
# where polling location strings are not the same
upload %>% 
  filter(Address_WEC == Address_VAN,
         `Polling Location_WEC` != `Polling Location_VAN` ) %>% 
  select(`Polling Location_WEC`,  
         `Polling Location_VAN`, 
         Address_WEC, Address_VAN, 
         City, 
         Zip, County) %>% 
  distinct() %>% 
  kablebox()
```

> We are not matching on Precinct Name. Where Precinct Names differ between the example addresses and VAN, we use the Precinct Name based on WEC's "Reporting Unit" (but we could do the reverse).


```{r}
upload %<>% 
  # take the State's name first 
  mutate(`Polling Location` = coalesce(`Polling Location_WEC` , 
                                       `Polling Location_VAN`),
         Address = Address_WEC) %>% 
  # select variable names from van
  select(any_of(names(van)), Address_WEC, Address_VAN,  PollingPlaceAddress) %>% 
  distinct()
```

### VAN polling place addresses not in WEC's file

#### Cases where the polling location address that is in VAN is not matching in WEC file

These polling places may have been deleted
```{r}
# polling addresses not in VAN? 
deleted <- van %>%  filter(!Address %in% upload$Address_WEC) %>% 
  select(County,  Address, `Polling Location`, `Precinct Name`, `Date Authenticated*`, `Precinct ID`) %>% 
  rename(Address_VAN = Address) %>%
  distinct() 

deleted %>% kablebox()
```

### WEC polling place addresses not in VAN

#### Cases where the polling location address is not matching in VAN 

These may be new polling places
```{r}
# polling addresses not in VAN? 
added <- upload %>% 
  filter(Address_WEC != Address_VAN | is.na(Address_VAN))  %>% 
  select(County, `Precinct Name`,Address_WEC, Address_VAN, `Polling Location`, `Date Authenticated*`, `Precinct ID`) %>% 
  distinct() 

added %>% kablebox()

deleted %<>% mutate(Issue = "Deleted address")
added %<>% mutate(Issue ="New address")

poll_ward_changes <- full_join(added, deleted)  

VANmissing <- read_csv(here::here("data", "VAN_missing_precincts.csv"))

poll_ward_changes %<>% anti_join(VANmissing %>% select(`Precinct Name`, `Address`)) %>% separate(`Precinct Name`, into = c("county", "Muni", "ward"), sep = " - ") %>% 
  select(-county)

# collapse
poll_address_changes <- poll_ward_changes %>%
  group_by(County, Muni, Address_VAN, Address_WEC) %>% 
  mutate(ward = ward %>% str_remove("WARD 0*") %>% str_c(collapse = ","),
         `Precinct ID` = `Precinct ID` %>% str_c(collapse = ",")) %>% 
  ungroup() %>%
  add_count(Muni) %>% 
  distinct() %>%
  mutate(ReportingUnit = paste(Muni, "WARD", ward)) %>% 
  select(-ward)

# arrange by number of issues per Muni
poll_address_changes %<>%
  arrange(-n, County, ReportingUnit) %>% 
  select(n, Issue, County, ReportingUnit, everything() )

poll_address_changes %>% kablebox()
```

Potentially added or deleted polling locations are automatically created [here](https://docs.google.com/spreadsheets/d/1cFJX0yGY91bzDpaR9ODwCeSCbzUrkWpRb1m7HwMKOA4/edit#gid=740092819). When we confirm that these are actual changes and not erroneous mismatches, they should be added to the [Polling Location Issue Tracker](https://docs.google.com/spreadsheets/d/1GYjifHWJ-N_Z3YEHMVWp58lA5d_emrzvgd4O951UTdU/edit#gid=120487606)

```{r, eval=FALSE}
library(googlesheets4)

ss <- googledrive::drive_get("Nov projected dem vote by ward") %>%
  gs4_get()

votes <- read_sheet(ss) %>% 
  transmute(`Precinct ID` = van_precinct_id,
            projected_dem_vote = projected_dem_vote) #%>% separate(van_precinct_name, into = c("County", "Muni", "ward"), sep = " - ")

#FIXME calculate diff in voters with a poll in their ward when we have wards of polling addresses

# votes %>% full_join(poll_ward_changes %>% select(`Precinct ID`, `Precinct Name`))

poll_address_changes <- poll_ward_changes %>% 
  left_join(votes) %>% 
  group_by(County, Muni, Address_VAN, Address_WEC) %>% 
  mutate(ward = ward %>% str_remove("WARD 0*") %>% str_c(collapse = ","),
         `Precinct ID` = `Precinct ID` %>% str_c(collapse = ","),
         projected_dem_vote = sum(projected_dem_vote)) %>% 
  ungroup() %>%
  add_count(Muni, name = "Wards Changing") %>% 
  distinct() %>%
  add_count(Muni, Issue) %>% 
  mutate(ReportingUnit = paste(Muni, "WARD", ward)) %>% 
  add_count(ReportingUnit) %>% 
  mutate(FLAG = ifelse(Issue == "Deleted address" & nn ==1, "!!!", "")) %>% 
  mutate(Issue = paste(n, FLAG, Issue)) %>% 
  select(-ward, -Muni, -n, - FLAG) 

# arrange by number of issues per Muni
poll_address_changes %<>%
  arrange(-`Wards Changing`, County, ReportingUnit) %>% 
  select(County, 
         Issue, ReportingUnit,
         `Wards Changing`,
         projected_dem_vote, 
         everything() )

poll_address_changes %>% kablebox()


#pull previously flagged issues and put new ones at the top
ss <- googledrive::drive_get("Polling Location Issue Tracker") %>%
  gs4_get() 

poll_address_changes_old <- sheets_read(ss, "Observed from WEC Data")

# note new and old
new_issues <- anti_join(poll_address_changes, poll_address_changes_old) %>% 
  mutate( date_added = Sys.Date() )

out <- full_join(poll_address_changes_old,  new_issues) %>% arrange(date_added)

# write to sheet
ss <- googledrive::drive_get("WI polling locations") %>%
  gs4_get() 


out %>%
  sheet_write(ss = ss,
              sheet = str_c("Issues ", Sys.Date() %>% as.character() ))
```


### All locations being updated in any way

(including location name changes and new addresses)

In old VAN, not in upload:
```{r}
anti_join(van, upload, by = c("County", "Address", "Polling Location")) %>% kablebox()
```

In new upload, not in old VAN:
```{r}
anti_join(upload, van, by = c("County", "Address", "Polling Location")) %>% kablebox()
```

# Save new polling location data 

```{r, eval=FALSE}
write_csv(upload, path = here("data", upload_file))

library(googlesheets4)



## drop "ghost wards missing from van"
ss <- googledrive::drive_get("WI polling locations") %>%
  gs4_get() 


# ghost wards 
ghosts <- read_sheet(ss, "VAN missing precincts")

ghosts %<>% select(`Precinct Name`, Address)

upload %<>% anti_join(ghosts)

upload %>%
  mutate(`Precinct ID` = `Precinct ID` %>% replace_na(0)) %>% 
  arrange(`Precinct ID`) %>% 
  sheet_write(ss = ss,
              sheet = paste("VAN Polling Locations", `Date Authenticated*`))
```

Here are the first few rows of the upload:
```{r}
upload %>% head() %>% kablebox()
```


## VAN-formatted combined elections.wi.gov + Google Civic polling locations:

CSV file: `r paste0("https://raw.githubusercontent.com/judgelord/vote/master/data/", upload_file) %>% str_replace_all(" ", "%20")`

Google Sheet: https://docs.google.com/spreadsheets/d/1cFJX0yGY91bzDpaR9ODwCeSCbzUrkWpRb1m7HwMKOA4

---

# Q & A
- Which fields from the VAN must be replaced exactly? If the state changes how `Polling Location` is described, do we want to replace it or have it fail to match with the `Precinct ID` from the VAN template? There are over 200 polling locations with slightly different names in the file on the state's website and the old VAN data.

>Precinct ID, Precinct Code, and Precinct Name are ones that should always be the same from VAN. If the state's name is slightly different for a polling location, if it's easier to just change it to that, that is fine. If the polling location is described differently, just replace it with what the state has. I think that answers your question, but let me know if it doesn't.

- There are over 200 polling location addresses in the State's 8-11-2020 xlsx file that I do not see in the VAN list. I only investigated the first one, 2100 Beaser Ave, Ashland, WI, and I am not clear if Bayfield Ward 12 is now voting there or not. Is there a simple explanation? If they need to be investigated, Can I anything with the data to make these cases easier to investigate? 

>These are most likely what we call ghost wards, where no one lives in. So we just exclude them from uploading to VAN since they don't have a VAN Precinct ID.

- The state's `Muni` is not the same as `City` in the VAN data for almost 2000 polling place addresses. For now, `Muni` is not being used in the merge. 
- If we are trying to align `City` and `Muni`, `City` in the VAN data does not contain "Town of"/"City of"/"Village of" but both the state's poll locations `Muni` xlsx file and the addresses file do. We are dropping the epithet, correct?

>This is probably because, in smaller munis, the actual address for a place has another muni in it. An example is the Town of Adams in Adams County. The actual address for the Town of Adams Town Hall has Friendship as the muni. So I would continue not using muni to match. We don't need to align the actual municipality name with the address municipality name. 

- If Google Civic has polling locations that the file on the state website does not, what should `Authenticating Authority` be? It is "Google Civic API" for now. 
- `Authority Contact Info` is "DPW- Fall Partisan Primary" in the VAN file. Do we want this to take the new file name or WEC's phone number?

>This is something that gets applied in bulk after I upload polling locations. So if it's easier for you to just not deal with those, that's fine. If it isn't, then yeah, Google Civic can be the authenticating authority and DPW-November Election as contact info

- `Location Description` in the VAN file contains poll hours. Do we want to merge this into the new upload or try to pull it in from Google Civic? Or both? Or neither?

> Keep location description as is with the polling hours. 

- I assume we are merging in `Precinct Code` from the VAN file on `Precinct ID`, correct? 

>Precinct ID is the primary key for uploading into VAN


- If a polling location is removed (i.e., it was in the previous VAN upload), but it is not in the new data, do we keep it in the upload sheet with some kind of a note, or do we just note on this data merge page that locations were removed?

> Don't need it noted in the file to upload but noted in the data merge page works. 

# TODO 

- split out reporting units from the state to see if wards move addresses (not just if there are new or deleted addresses)

# Notes about data for VAN

1. The Addresses in the VAN data capitalize oddly--"1St" "2Nd" "3Rd" etc.--I made all strings all caps.

# Additional Google Civic API Contents

For reference, these values may or may not be returned: 
```
"pollingLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "earlyVoteSites": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "dropOffLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
```