---
title: "Polling Place Changes in WI"
subtitle: 
author:
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE, 
                      #fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)

library(httr)
library(jsonlite)
library(tidyverse)
library(magrittr)
library(here)
library(knitr)
library(kableExtra)
library(rvest)

kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```

> Note: Below are **election day** polling places. For early voting (October 20-30) and ballot drop box locations, go to [iwillvote.com](https://iwillvote.com/locate/?state=WI)

# Polling Places from the Wisconsin Elections Commission

```{r}
# UPDATE THIS IF A NEW FILE IS POSTED! 
file <- "2020 General Election PPL RU List.xlsx"

official_locations <- readxl::read_xlsx(here::here("data", file))

date <- "8-11-2020"
```

This analysis uses the "`r file`" data posted on `r date`. Here are the current files posted on https://elections.wi.gov/node/6527:

```{r}
# polling locations posted on website
html <- read_html("https://elections.wi.gov/node/6527") %>%
  html_nodes("a") 

files <- tibble(file = html_text(html),
                link = html_attr(html, "href")) %>% 
  filter(str_detect(file, "xls|csv")) 
                
kable(files) %>% 
  kable_styling() %>% 
  scroll_box()
``` 

---

# Changes from the August Primary Election

```{r}
old_locations <- readxl::read_xlsx(here::here("data", "PPL 2020 Partisan Primary 2.xlsx")) %>% 
  select(County, Muni, ReportingUnit, PollingPlaceName, PollingPlaceAddress, Longitude, Latitude)

# a function to split out wards
split <- function(ward){
  
  min = str_extract(ward, "[0-9]*") %>% 
           as.numeric()
  
  max = str_extract(ward, "-[0-9]*") %>% 
           as.numeric() %>% abs()
  
  full_seq(c(min,
             max),
           1) %>%
    str_c(collapse = ",") 
}

# split out wards
split_wards <- . %>%
  mutate_if(is.character, str_to_upper) %>% 
  mutate(ward = str_extract(ReportingUnit, "[1-9].*") %>% 
           # split on commas in wards
           str_split(",")) %>% 
  unnest(ward) %>% 
  # remove any letters
  mutate(ward = ward %>% str_remove("[A-z]")) %>% 
  # make a sequence of wards
  mutate(ward_seq = map_chr(ward, possibly(split, otherwise = NA)),
         ward = coalesce(ward_seq, ward) %>% 
           str_split(",")) %>% 
  unnest(ward) %>% 
  mutate_if(is.character, str_to_title) %>%
  mutate(County = str_remove(County, " County")) %>%
  mutate(ReportingUnit = paste("Ward ", ward)) %>% 
  ungroup() %>% 
  select(-ward_seq)


# apply function to old and new polling locations
official <- official_locations %>% split_wards()
old <- old_locations %>% split_wards()

# diff between old and new polling locations
diff <- anti_join(official, old, 
                  by = c("County", "Muni", "ReportingUnit", 
                         "ward", "PollingPlaceAddress")) %>%
  left_join(old, 
            by = c("County", "Muni", "ReportingUnit", "ward"),
            suffix = c(".General", ".Primary"))

# simplify 
diff %<>% mutate(General = str_c(PollingPlaceName.General, 
                                 PollingPlaceAddress.General, 
                                 sep = ", "),
                 Primary = str_c(PollingPlaceName.Primary, 
                                 PollingPlaceAddress.Primary, 
                                 sep = ", "))
```




```{r map_change}
wi <- map_data("county", "wisconsin") %>% mutate(County = str_to_title(subregion))

diff %<>% 
  group_by(County)%>% 
  mutate(changes = length(unique(PollingPlaceAddress.General)),
         net = changes - length(unique(PollingPlaceAddress.Primary))) 
wi  %<>% 
  # add counts 
  left_join(diff %>% 
              select(County, Muni, PollingPlaceAddress.General, changes, net) ) %>% 
  mutate(changes = changes %>% replace_na(0)) %>% 
  # add points 
  full_join(diff %>% mutate(group = NA)) %>% 
  mutate(changes = changes %>% replace_na(0),
         net = net %>% replace_na(0)) 

# changes 
wi %>% 
  group_by(subregion) %>%
  # Find the center of each subregion
  mutate(center_lat = mean(range(lat) ),
         center_long = mean(range(long) ) ) %>%
ggplot( aes(x = long, y = lat, group = group, fill = changes) ) +
  geom_polygon(color = "white") +
  geom_point(aes(x = Longitude.Primary, 
                 y = Latitude.Primary), 
             color = "Red", alpha = .5) +
  geom_point(aes(x = Longitude.General, 
                 y = Latitude.General), 
             color = "Green", shape = "+", size = 5, alpha = .5) + 
  geom_text( aes(x = center_long, y = center_lat, label = County), size = 2, angle = 45, check_overlap = T, color = "white") +
  theme_void() + 
  labs(fill = "Polling Place\nAddresses Changed\nper County")
```

New polling locations are green. Removed locations are red. Brighter points mean that more Wards moved to or from that polling location. 

The geo data have a few errors, but these points are mostly correct.

```{r map_net}
# net 
wi %>% 
  group_by(subregion) %>%
  # Find the center of each subregion
  mutate(center_lat = mean(range(lat) ),
         center_long = mean(range(long) ) ) %>%
ggplot( aes(x = long, y = lat, group = group, fill = net) ) +
  geom_polygon(color = "white") +
  geom_point(aes(x = Longitude.Primary, 
                 y = Latitude.Primary), 
             color = "Red", alpha = .5) +
  geom_point(aes(x = Longitude.General, 
                 y = Latitude.General), 
             color = "Green", shape = "+", size = 5, alpha = .5) + 
  geom_text( aes(x = center_long, y = center_lat, label = County), 
             size = 2, angle = 45, check_overlap = T, color = "white") +
  theme_void() + 
  labs(fill = "Net Polling Places\nAdded per County")
```

Green Bay (Brown County) and Burlington (Racine County) now have *fewer* polling places than they did for the August Primary election.  



```{r}
# table of changes 
diff %>% 
  ungroup( ) %>%
  group_by(County, Muni, General, Primary) %>% 
  summarise(Ward = paste(County, Muni, "Ward", 
                         paste(ward, collapse = ", "),
                         sep = ", ")) %>%
  ungroup() %>% 
  select(Ward, Primary, General) %>%  
  distinct() %>% 
  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "600px")
```

```{r map_county}
# a function to plot each county
countymap <- function(county){
p <- wi %>% 
  filter(County == county) %>% 
ggplot( ) +
  geom_polygon(aes(x = long, y = lat, group = group)) +
  geom_point(aes(x = Longitude.Primary, 
                 y = Latitude.Primary), 
             color = "Red", alpha = .5) +
  geom_point(aes(x = Longitude.General, 
                 y = Latitude.General), 
             color = "Green", shape = "+", size = 5, alpha = .5) + 
  geom_text( aes(x = Longitude.Primary, 
                 y = Latitude.Primary,
                 label = PollingPlaceName.Primary), 
             color = "Red", alpha = .5, 
             angle = 45, check_overlap = T) +
    geom_text( aes(x = Longitude.General, 
                 y = Latitude.General,
                 label = PollingPlaceName.General), 
             color = "Green", alpha = .5, 
             angle = 45, check_overlap = T) +
  geom_segment(aes(x = Longitude.Primary,
                   y = Latitude.Primary,
                   xend = Longitude.General, 
                   yend = Latitude.General), 
                #arrow = arrow(length = unit(0.1, "cm")),
               color = "green") + 
  geom_text( aes(x = Longitude.General, 
                 y = Latitude.General,
                 label = PollingPlaceName.General), 
             color = "Green", alpha = .5, 
             angle = 45, check_overlap = T) +
  theme_void() + 
  labs(title = paste(county, "County"))

return(p)
}

counties <- map(diff %>%
      filter(net != 0 | changes > 5) %>%
      .$County %>% unique(),
    countymap)

counties
```

---

# Polling locations via Google Civic's API

(Actually, it goes Wisconsin Elections Commission --> Voter Information Project --> Google Civic's API)

To check that Google Civic is providing the same information that is posted on the Election Commission's website, we can look up polling locations for addresses in each  voting precinct using the Google Civic API.

```{r}
#########################################################
# one address per precinct
addresses <- read_csv(here::here("data", "wisconsin_example_addresses.csv"))

addresses %<>% mutate_if(is.character, str_to_upper)

addresses %<>% arrange(county_name)

# FIXME DELETE THIS WHEN API IS WORKING
addresses %<>% head(2)
# /FIXME
```



Polling locations are sometimes (in the days or weeks before an election) available via the Google Civic API, which uses data collected from states by the Voter Information Project, which notifies users when it posts data here: https://groups.google.com/g/vip-community

Notes about the Google Civic API (see the relevant documentation [here](https://developers.google.com/civic-information/docs/v2/elections/voterInfoQuery))  
- create an API key here: https://console.developers.google.com/apis/credentials  
- remember to authorize you API key  
- API calls should not need `electionId`; docs say it is optional  
- `pollingLocations` will not always be returned  
- may return a great deal more than polling locations (see Additional Google Civic API Contents below)
- may return more than one polling place per input address as a list 



```{r}
# make full street address
addresses %<>% 
  mutate(voter_address = paste(voting_street_address,
                               voting_city, 
                               state_code))

addresses$voter_address[1]
```

Here is the polling place from the Google Civic API for the above address:
```{r}
# need to get AND authorize an api key
source(here::here("api_key.R"))

# A function to get poll addresses
get_poll <- function(voter_address){
  
  voter_address %<>% 
    str_replace_all(" ", "%20") %>%
    str_remove_all(",")
  
  url <- str_c("https://civicinfo.googleapis.com/civicinfo/v2/voterinfo?address=", 
               voter_address, 
               "&returnAllAvailableData=true&key=", 
               api_key)
  
  
  poll <- GET(url) %>% 
    .$content %>% 
    rawToChar() %>% 
    fromJSON() %>%
    .$pollingLocations %>% # note: there is a lot more than this
    .$address %>% 
    {ifelse(is.null(.), "NULL", .)}
  
  return(poll)
}


# test 
get_poll(addresses$voter_address[1])

# apply function to addresses
api_locations <- addresses %>% 
  # "NULL" if NULL, "error" if any error
  mutate(PollingPlaceAddress = voter_address %>% map(possibly(get_poll, "error"))) 

# expand if there is more then one polling place per address
api_locations %<>% unnest(PollingPlaceAddress)

api_locations %>% select(voting_street_address, precinct_name, precinct_id, PollingPlaceAddress) %>%  kablebox()
```

### API results not in the state's xlsx file

```{r}
api_locations %>% 
  filter(PollingPlaceAddress != "NULL") %>% # filter out API errors
  anti_join(official_locations) %>% # filter out locations already posted
  select(precinct_name, precinct_id, PollingPlaceAddress) %>%
  kablebox()
```

### Cases where Google Civic's API did not return the same address when given polling locations

> Until the API is working, it is all of them. I'm just showing a few for now.

```{r}
official_locations %>% 
  head() %>% #FIXME WHEN THE API IS WORKING
  mutate(PollingPlaceAddress_API = map(PollingPlaceAddress, get_poll)) %>% # get polling places for polling places' addresses
  filter(PollingPlaceAddress_API != PollingPlaceAddress) %>% # mismatches
  select(Muni, ReportingUnit, PollingPlaceName, PollingPlaceAddress, PollingPlaceAddress_API) %>%
  kablebox()
```

### Polling place addresses not returned by Google Civic's API

> Until the API is working, it is all of them. I'm just showing a few for now.

```{r}
official_locations %>% 
  head() %>% #FIXME WHEN THE API IS WORKING
  anti_join(api_locations) %>% # remove ones matched with API
  select(Muni, ReportingUnit, PollingPlaceName, PollingPlaceAddress) %>% 
  kablebox()
```

---


```{r}
# align official and Google Civic API results
api_locations %<>% rename(County = county_name)

# reformat county
official_locations %<>% 
  mutate_if(is.character, str_to_title) %>% 
  mutate(County = str_remove(County, " COUNTY") )


locations <- full_join(official_locations, api_locations)
```







## Additional Google Civic API Contents

For reference, these values may or may not be returned: 
```
"pollingLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "earlyVoteSites": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
  "dropOffLocations": [
    {
      "id": string,
      "address": {
        "locationName": string,
        "line1": string,
        "line2": string,
        "line3": string,
        "city": string,
        "state": string,
        "zip": string
      },
      "notes": string,
      "pollingHours": string,
      "name": string,
      "voterServices": string,
      "startDate": string,
      "endDate": string,
      "latitude": double,
      "longitude": double,
      "sources": [
        {
          "name": string,
          "official": boolean
        }
      ]
    }
  ],
```